---
title: "Comparing Two Portfolios in Real Time"
author: '<span class="meta">Lemuel Kumarga</span>'
date: '<span class="meta">Dec 2019</span>'
always_allow_html: yes
knit: (function(inputFile, encoding) { source("shared/knit.R"); knitPPT(inputFile, encoding) })
---

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
packages <- c("knitr","highcharter","DiagrammeR")
tmp <- lapply(packages, library, character.only = TRUE)

# Set images globally
opts_chunk$set(fig.align='center', fig.height = 5, fig.width = 8)

# Load default settings for R Markdown -- see file for more details
source("shared/defaults.R")
# Load some helper functions
source("shared/helper.R")

options(stringsAsFactors = FALSE)

data_dir <- "data/"
output_dir <- "output/"
```

<!-- Additional JS CDN for this project -->
<script src="https://cdnjs.cloudflare.com/ajax/libs/moment.js/2.24.0/moment.min.js"></script>
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">

## Problem Description

With the rise of internet productivity tools such as email, text messaging and social media, companies are required to be nimble and act within a couple of hours.

Fortunately, advancements in hardware have allowed us to parse information in real time, providing wisdom in making these split-second decisions.

In this portfolio, we will be building a simple real time tracker through AWS and visualizing it in Highcharts.

## A Simple Use Case

In this study, we will be comparing the performance of two exchange traded funds, <a href="https://investor.vanguard.com/etf/profile/VTI" target="_blank">VTI</a> and <a href="https://investor.vanguard.com/etf/profile/VXUS" target="_blank">VXUS</a>.

Both track the economy of United States, and the rest of the world respectively.

Performances are defined by their <a data-toggle="popover" title="Daily Returns" data-content="% gain when you sold the stock bought 24 hours ago">daily returns</a>: 
<br>

$$ \frac{\text{Current Price} - \text{Price 24 Hours Ago}}{\text{Price 24 Hours Ago}}$$

The returns are measured in basis points ( 1 / 100th percent).

## The Real Time Tracker

```{r echo=FALSE, warning=FALSE}
hcharts_opts = hcharts_ppt()
hcharts_opts[['data']] = list(csvURL='https://j5ah51nweg.execute-api.ca-central-1.amazonaws.com/api/returns',
                              enablePolling =TRUE,
                              dataRefreshRate = 5)

# Modify some defaults
hcharts_opts$tooltip$headerFormat = NULL

highchart(hc_opts=hcharts_opts, type='stock') %>%
  hc_navigator(enabled=FALSE) %>%
  hc_rangeSelector(enabled=FALSE) %>%
  hc_scrollbar(enabled=FALSE) %>%
  hc_title(text="VTI vs VXUS Performance") %>%
  hc_xAxis(tickPositioner=JS("function () {
                          			var positions = []
                          
                                      // Find UTC timezone offset to New York for DST adjustment
                                      var cur_d = new Date()
                                      var utc_d = new Date(cur_d.toLocaleString('en-US', { timeZone: 'UTC' }))
                                      var ny_d = new Date(cur_d.toLocaleString('en-US', { timeZone: 'America/New_York' }))
                                      var tz_offset = (ny_d.getTime() - utc_d.getTime())/1000
                          
                                      for (var tick = Math.ceil(this.dataMin / 60000) * 60000; 
                                      	 tick <= this.dataMax - 30000; tick += 60000) {
                                      	
                                      	var d=new Date(0)
                                      	// Default tick value is in local timezone
                                      	// Convert the local timezone to utc then 
                                      	// move it into NY timezone
                                      	var ny_time = tick + moment().zone() *60 * 1000 + tz_offset * 1000
                                      	d.setTime(ny_time)
                          
                                      	// Skip Non-Trading Hours
                                      	// If we are on Friday after 4:00pm, skip for two days
                                      	if (d.getUTCDay() == 5 && d.getUTCHours() >= 16) {
                                      		tick += 2 * 24 * 3600 * 1000
                                      	}
                                      	// If we are after 4pm, skip until 9:30am
                                      	if (d.getUTCHours() >= 16) {
                                      		tick += 17.5 * 3600 * 1000 
                                      		continue
                                      	}	
                          
                                      	positions.push(tick);
                                      }
                          
                                      positions.push(this.dataMax)
                          
                                      return positions
                            }"),
              labels=list(useHTML = FALSE,
                       align='right',
                       formatter=JS("function() {
                                        var d = new Date(0);
                                        // By default date string to epoch conversion in Highcharts
                                        // takes into account timezone - hence this.value is
                                        // already in user's timezone
                                        d.setTime(this.value);
                        
                                        var hours = d.getUTCHours();
                                        if (hours < 10) hours = '0' + hours;
                        
                                        var minutes = d.getUTCMinutes();
                                        if (minutes < 10) minutes = '0' + minutes;
                        
                                        var seconds = d.getUTCSeconds();
                                        if (seconds == 0) seconds = '';
                                        else if (seconds < 10) seconds = '0' + seconds;
                                        return hours + ':' + minutes + (seconds == '' ? '' : ':' + seconds)
                                    }")),
           tickWidth = 0.5,
           crosshair=TRUE) %>%
  hc_yAxis(title=list(margin = 0,
                      text="Daily Returns (bps)"),
           labels=list(useHTML=FALSE,
                       verticalAlign= 'middle',
                       formatter=JS("function () {
                            var label = this.axis.defaultLabelFormatter.call(this);
                            return Highcharts.numberFormat(this.value * 10000.,0);
                        }")),
           gridLineWidth = 1,
           tickWidth = 0.5,
           opposite=FALSE) %>%
  hc_legend(useHTML=FALSE,
            align='right',
            enabled=TRUE) %>%
  hc_plotOptions(series=list(
    dataLabels=list(
      enabled=TRUE,
      padding=5,
      align='right',
      style = list(
        fontFamily = `@f`,
        fontSize = '0.8em',
        fontWeight = 'normal'
      ),
      formatter=JS("function(){
                    var isLast = false;
                    
                    if(this.series.data.length == 0) return '';
                    if(this.point.x == this.series.data[this.series.data.length - 1].x) isLast = true;

                    if (isLast) {
                        return '<span style=\"color:' + this.series.color + '\">' + this.series.name + ': ' + (this.y * 10000.).toFixed(2) + ' bps</span>';
                    }
                }")
    )
  )) %>%
  hc_tooltip(useHTML=FALSE,
             pointFormatter=JS("function() {
             return this.series.name + ': ' +
                    (this.y * 10000.).toFixed(2) + ' bps';
              }"),
             shared=TRUE) %>%
  hc_series(list(color=`@c`(1)),
            list(color=`@c`(2)))

```


# Building The Data Pipeline

## A Schematic Overview
```{r echo=FALSE}
grViz(paste0("
  digraph pipeline {

    # Default Specs
    graph [compound = true, nodesep = .5, ranksep = .25]
    node [fontname = '",`@f`,"', fontsize = 14, fontcolor = '",`@c`(bg),"', penwidth=0, style=filled]
    edge [fontname = '",`@f`,"', color='",`@c`(ltxt),"', fontcolor = '",`@c`(ltxt),"']
    
    # Writing Nodes
    DB [label = 'Database', color = '",`@c`(txt,0.1),"', fontcolor = '",`@c`(txt),"', shape= rectangle, width = 5]
    
    UT_Script [label = 'updateTable\nScript', color = '",`@c`(1),"', shape=diamond]
    AV_API [label = 'Stocks API', color = '",`@c`(txt),"', shape=rectangle]
    { rank = same; UT_Script, AV_API}

    AV_API -> UT_Script -> DB [ dir=back, color='",`@c`(1),"' ]
    AV_API -> UT_Script -> DB [ color='",`@c`(1),"' ]
    
    # Reading Nodes
    Highcharts [label = 'Tracker', color = '",`@c`(bg),"', fontcolor = '",`@c`(txt),"', shape=plaintext]
    AWS_API [label = 'AWS API', color = '",`@c`(txt),"', shape=plaintext]
    GR_Script [label = 'getReturns\nScript', color = '",`@c`(2),"', shape = diamond]
    { rank = same; AWS_API, GR_Script}
    
    DB -> GR_Script [dir=back, color='",`@c`(2),"']
    GR_Script -> AWS_API -> Highcharts [dir=back, color='",`@c`(2),"']
    
    DB -> GR_Script [color='",`@c`(2),"']
    GR_Script -> AWS_API -> Highcharts [color='",`@c`(2),"']
    
  }
"))
```

<span style="font-size: 12pt"><span style="color:`r '@c'(1)`"><i class="fa fa-arrow-right"></i></span>: Writing&#8195;&#8195;&#8195;&#8195;&#8195;&#8195;<span style="color:`r '@c'(2)`"><i class="fa fa-arrow-right"></i></span>: Reading</span>

## Database & Structure

We choose <a href="https://aws.amazon.com/dynamodb/" target="_blank">Amazon DynamoDB</a> as the database since it provides sufficient free storage, writing and reading capacities.

The database is organized in the following format:

<b class="color-1-text">Ticker</b>: The ID of the index fund (VTI)<br>
<b class="color-1-text">Time</b>: Time of day (09:30:30)<br>
<b class="color-1-text">Last Price</b>: Last recorded prices(151.00)<br>
<b class="color-1-text">Last Returns</b>: Last calculated returns (0.00005)<br>
<b class="color-1-text">Last TS</b>: Last modified timestamp (2019-12-12 09:30:30)<br>

## Writing the Data

There are 3 main components involved in writing:

- <b class="color-1-text">Stock API</b>: This API provides real-time stock/fund prices. In this project, we use <a href="https://www.alphavantage.co" target="_blank">Alpha Vantage</a> due to its free tier offerings.
- <b class="color-1-text">Database</b>: Where all information is stored.
- <b class="color-1-text"><a href="updateTable.zip" download>updateTable Script</a></b>: This Python script extracts raw information from the other 2 components, transforms it and writes it back into the database.

## The Writing Pipeline

```{r echo=FALSE}
grViz(paste0("
  digraph writePipeline {

    # Default Specs
    graph [rankdir = LR, compound = true, nodesep = .5, ranksep = .25]
    node [fontname = '",`@f`,"', fontsize = 12, penwidth=0, style=filled, shape=rectangle]
    edge [fontname = '",`@f`,"', fontsize = 12, color='",`@c`(ltxt),"', fontcolor = '",`@c`(ltxt),"']
    
    node [color='",`@c`(txt),"', fontcolor = '",`@c`(bg),"' ]
    AP_1 [ label = 'API' ]
    
    node [color='",`@c`(1),"', shape=diamond]
    SC_1 [ label = 'Script' ]
    SC_2 [ label = 'Script' ]
    SC_3 [ label = 'Script' ]
    SC_4 [ label = 'Script' ]
    {rank = same; SC_1, SC_2, SC_3, SC_4 }
    
    node [color='",`@c`(txt,0.1),"', fontcolor = '",`@c`(txt),"', shape=rectangle]
    DB_1 [label = 'Database']
    DB_2 [label = 'Database']
    {rank = same; DB_1, DB_2 }
    
    AP_1 -> SC_1 [dir=back, label='Ping\nevery 30 min'] 
    AP_1 -> SC_2 [ label = 'Return\nlatest prices' ]
    SC_2 -> DB_1 [ label = 'Query last\nstored prices' ]
    SC_3 -> DB_1 [ dir=back, label = 'Return prices' ]
    SC_3 -> SC_4 [ label = 'Calculate          \\rreturns          \\r', labeldistance = 50 ]
    SC_4 -> DB_2 [ label = 'Write returns,\ntimestamp\nand latest\nprices' ]
  }
"))
```

## Reading The Data { .slide-col2 }

There are 4 main components in the reading process:

- <b class="color-2-text">Tracker</b>: The tracker pings every 5 seconds to retrieve the latest returns.
- <b class="color-2-text"><a href="https://aws.amazon.com/api-gateway/">Amazon API</a></b>: This API acts as the gateway into the back-end ecosystem. It receives the pings and returns information to the tracker.
- <b class="color-2-text">Database</b>: Where all the information is stored.
- <b class="color-2-text"><a href="getReturns.py" download>getReturns Script</a></b>: This Python script queries the latest returns from the database and formats them in a digestible manner.

## The Reading Pipeline { .slide-col2 }

```{r echo=FALSE}
grViz(paste0("
  digraph readPipeline {

    # Default Specs
    graph [rankdir = LR, compound = true, nodesep = .5, ranksep = .25]
    node [fontname = '",`@f`,"', fontsize = 12, penwidth=0, style=filled, shape=rectangle]
    edge [fontname = '",`@f`,"', fontsize = 12, color='",`@c`(ltxt),"', fontcolor = '",`@c`(ltxt),"']
    
    
    node [color='",`@c`(txt),"', fontcolor = '",`@c`(bg),"']
    AP_1 [ label = 'API' ]
    AP_2 [ label = 'API' ]
    
    node [color='",`@c`(txt,0.1),"', fontcolor = '",`@c`(txt),"']
    DB_1 [label = 'Database']
    {rank = same; AP_1, AP_2, DB_1 }   
    
    node [color='",`@c`(bg),"']
    TK_1 [ label = 'Tracker' ]
    TK_2 [ label = 'Tracker' ]
    {rank = same; TK_1, TK_2} 

    node [color='",`@c`(2),"', fontcolor = '",`@c`(bg),"', shape=diamond]
    SC_1 [ label = 'Script' ]
    SC_2 [ label = 'Script' ]
    SC_3 [ label = 'Script' ]
    {rank = same; SC_1, SC_2, SC_3 }
    
    
    TK_1 -> AP_1 [label='Ping every\n5 seconds'] 
    AP_1 -> SC_1 [label='Request\nreturns\n']
    DB_1 -> SC_1 [dir=back,label='Query\nreturns']
    DB_1 -> SC_2 [label='Return\nreturns']
    SC_2 -> SC_3 [label='Reformat      \\rdata      \\r']
    AP_2 -> SC_3 [dir=back, label='Send\nreturns']
    TK_2 -> AP_2 [dir=back, label='Return\nreturns']
    AP_1 -> DB_1 [color='",`@c`(bg),"']
  }
"))
```

# Limitations

## The Database

Under the current process, past prices of funds are replaced with new ones. 

This approach was chosen so that DynamoDB storage would always fall within the free tier. If costs were not an issue, we should refrain from deleting for any future use.

Alternatively, we could also utilize <a href="https://aws.amazon.com/s3/" target="_blank">S3</a> and <a href="https://aws.amazon.com/ec2/" target="_blank">EC2</a> since <a href="https://blog.yugabyte.com/dynamodb-pricing-calculator-expensive-vs-alternatives/" target="_blank">DynamoDB's pricing may increase significantly with scale</a>.

## The 30 Minute Lag

The tracker is not purely real-time since there is a 30 minute lag.

The lag is due to <a href="https://www.alphavantage.co" target="_blank">Alpha Vantage</a>'s free tier limitations.

By purchasing the premium tier, we could potentially reduce the lag to 1 minute.
