---
title: "Comparing Two Portfolios in Real Time"
author: '<span class="meta">Lemuel Kumarga</span>'
date: '<span class="meta">Dec 2019</span>'
always_allow_html: yes
knit: (function(inputFile, encoding) { source("shared/knit.R"); knitPPT(inputFile, encoding) })
---

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
packages <- c("knitr","highcharter")
tmp <- lapply(packages, library, character.only = TRUE)

# Set images globally
opts_chunk$set(fig.align='center', fig.height = 5, fig.width = 8)

# Load default settings for R Markdown -- see file for more details
source("shared/defaults.R")
# Load some helper functions
source("shared/helper.R")

options(stringsAsFactors = FALSE)

data_dir <- "data/"
output_dir <- "output/"
```


## Problem Description

With the rise of the internet productivity tools such as email, text messaging and social media, more companies are required to be nimble and act within a couple of hours.

Fortunately, advancements in hardware have allowed us to parse information in real time, providing wisdom in making these split-second decisions.

In this portfolio, we will be building a simple real time tracker through AWS and visualizing it in Highcharts.

## A Simple Use Case

In this study, we will be comparing the performance of two exchange traded funds, <a href="https://investor.vanguard.com/etf/profile/VTI" target="_blank">VTI</a> and <a href="https://investor.vanguard.com/etf/profile/VXUS" target="_blank">VXUS</a>.

Both track the economy of United States, and the rest of the world respectively.

Performances are defined by their <a data-toggle="popover" title="Daily Returns" data-content="% gain when you sold the stock bought 24 hours ago">daily returns</a>: 
<br>

$$ \frac{\text{Current Price} - \text{Price 24 Hours Ago}}{\text{Price 24 Hours Ago}}$$

The returns are measured in basis points ( 1 / 100th percent).

## The Real Time Tracker

```{r echo=FALSE, warning=FALSE}
hcharts_opts = hcharts_ppt()
hcharts_opts[['data']] = list(csvURL='https://j5ah51nweg.execute-api.ca-central-1.amazonaws.com/api/returns',
                              enablePolling =TRUE,
                              dataRefreshRate = 5)

# Modify some defaults
hcharts_opts$tooltip$headerFormat = NULL

highchart(hc_opts=hcharts_opts, type='stock') %>%
  hc_navigator(enabled=FALSE) %>%
  hc_rangeSelector(enabled=FALSE) %>%
  hc_scrollbar(enabled=FALSE) %>%
  hc_title(text="VTI vs VXUS Performance") %>%
  hc_xAxis(tickPositioner=JS("function () {
                          			var positions = []
                          
                                      // Find UTC timezone offset to New York for DST adjustment
                                      var cur_d = new Date()
                                      var utc_d = new Date(cur_d.toLocaleString('en-US', { timeZone: 'UTC' }))
                                      var ny_d = new Date(cur_d.toLocaleString('en-US', { timeZone: 'America/New_York' }))
                                      var tz_offset = (ny_d.getTime() - utc_d.getTime())/1000
                          
                                      for (var tick = Math.ceil(this.dataMin / 60000) * 60000; 
                                      	 tick <= this.dataMax - 30000; tick += 60000) {
                                      	
                                      	var d=new Date(0);
                                      	// Default tick value is in local timezone
                                      	// Convert the local timezone to utc then 
                                      	// move it into NY timezone
                                      	d.setUTCSeconds(tick/1000 + d.getTimezoneOffset()*60 + tz_offset);
                          
                                      	// Skip Non-Trading Hours
                                      	// If we are on Friday after 4:00pm, skip for two days
                                      	if (d.getUTCDay() == 5 && d.getUTCHours() >= 16) {
                                      		tick += 2 * 24 * 3600 * 1000
                                      	}
                                      	// If we are after 4pm, skip until 9:30am
                                      	if (d.getUTCHours() >= 16) {
                                      		tick += 17.5 * 3600 * 1000 
                                      		continue
                                      	}	
                          
                                      	positions.push(tick);
                                      }
                          
                                      positions.push(this.dataMax)
                          
                                      return positions
                            }"),
              labels=list(useHTML = FALSE,
                       align='right',
                       formatter=JS("function() {
                                        var d = new Date(0);
                                        d.setUTCSeconds(this.value/1000 + d.getTimezoneOffset()*60);
                        
                                        var hours = d.getHours();
                                        if (hours < 10) hours = '0' + hours;
                        
                                        var minutes = d.getMinutes();
                                        if (minutes < 10) minutes = '0' + minutes;
                        
                                        var seconds = d.getSeconds();
                                        if (seconds == 0) seconds = '';
                                        else if (seconds < 10) seconds = '0' + seconds;
                                        return hours + ':' + minutes + (seconds == '' ? '' : ':' + seconds)
                                    }")),
           tickWidth = 0.5,
           crosshair=TRUE) %>%
  hc_yAxis(title=list(margin = 0,
                      text="Daily Returns (bps)"),
           labels=list(useHTML=FALSE,
                       verticalAlign= 'middle',
                       formatter=JS("function () {
                            var label = this.axis.defaultLabelFormatter.call(this);
                            return Highcharts.numberFormat(this.value * 10000.,0);
                        }")),
           gridLineWidth = 1,
           tickWidth = 0.5,
           opposite=FALSE) %>%
  hc_legend(useHTML=FALSE,
            align='right',
            enabled=TRUE) %>%
  hc_plotOptions(series=list(
    dataLabels=list(
      enabled=TRUE,
      padding=5,
      align='right',
      style = list(
        fontFamily = `@f`,
        fontSize = '0.8em',
        fontWeight = 'normal'
      ),
      formatter=JS("function(){
                    var isLast = false;
                    
                    if(this.series.data.length == 0) return '';
                    if(this.point.x == this.series.data[this.series.data.length - 1].x) isLast = true;

                    if (isLast) {
                        return '<span style=\"color:' + this.series.color + '\">' + this.series.name + ': ' + (this.y * 10000.).toFixed(2) + ' bps</span>';
                    }
                }")
    )
  )) %>%
  hc_tooltip(useHTML=FALSE,
             pointFormatter=JS("function() {
             return this.series.name + ': ' +
                    (this.y * 10000.).toFixed(2) + ' bps';
              }"),
             shared=TRUE) %>%
  hc_series(list(color=`@c`(1)),
            list(color=`@c`(2)))

```


# Building The Data Pipeline

## A Brief Overview

## The AWS DynamoDB Table

## Writing The Data

## Reading The Data

# Takeaways

## Conclusion

## Limitations

## Dark Deck { .slide-invert }

- <a data-toggle="popover" title="Lorem Ipsum" data-content="Lorem Ipsum is simply dummy text of the printing and typesetting industry. Data obtained from https://www.lipsum.com/.">Hover Point</a>
- Point 1
- Point 2
- Point 3

## Incremental Deck

>- <a data-toggle="popover" title="Lorem Ipsum" data-content="Lorem Ipsum is simply dummy text of the printing and typesetting industry. Data obtained from https://www.lipsum.com/.">Hover Point</a>
>- Point 1
>- Point 2
>- Point 3
